Instantiating Bunched Dependent Type Theory for Monoidal Classifying Toposes

### 2025-01-21 paper.tex
Assume less familiarity with Schopp's work.

\begin{abstract}
    I provide a simple condition on a geometric theory \(\T\),
    which suffices to interpret the bunched type theory \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\) of \cite{schopp} into the classifying topos \(\Set[\T]\).
\end{abstract}

@@

To prove the main theorem, begin by recalling some results from \cite{schopp}.

\begin{lemma}
    The type theory \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\) can be modeled in any \\
    \((*, 1, \Sigma, \Pi, \Pi^*)\)-type-category.
\end{lemma}

\begin{proof}
    \cite{schopp}, chapter 6.
\end{proof}

\begin{lemma}
    Let \(\B\) be a topos equipped with a strict affine symmetric monoidal closed structure \((*,\wand)\),
    such that \(- * A\) preserves pullbacks for all \(A\).
    Further suppose we have a canonical choice of pullback for every span in \(\B\).
    Then the family fibration over \(\B\) is a \((*, 1, \Sigma, \Pi, \Pi^*)\)-type-category.
\end{lemma}

\begin{proof}
    \cite{schopp}, section 3.5. Since we are in a topos, the condition that the monomorphism \(A * B \hookrightarrow A \times B\) is strong is trivial.
\end{proof}

Next, specialize to sheaf toposes.

@@

Finally, set the site to a syntactic site, to understand classifying toposes.

@@

    \pagebreak

    The syntactic category is guaranteed to be finitely complete, with a canonical construction for pullbacks.
    Specifically, the empty context is terminal,
    and the pullback of the span \(\Gamma \xrightarrow{f} \Xi \xleftarrow{g} \Delta\) is the context \((\gamma : \Gamma, \delta : \Delta, f\gamma = g\delta)\).

>>>>

\begin{abstract}
    In their PhD thesis, Sch\"opp introduces a type theory designed to capture the monoidal structure in the Schanuel topos.
    I provide a simple condition that can be used to verify whether the type theory interprets into a classifying topos of your choice.
\end{abstract}

@@

To prove the main theorem, we begin by recalling a result from \cite{schopp}.

\begin{lemma}
    Let \(\B\) be a topos equipped with a strict affine symmetric monoidal closed structure \((*,\wand)\),
    such that \(- * A\) preserves pullbacks for all \(A\).
    Further suppose we have a canonical choice of pullback for every span in \(\B\).
    Then the type theory \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\) interprets into \(\B\).
\end{lemma}

\begin{proof}
    Chapter 6 of \cite{schopp} shows that we can interpret the type theory in any ``\((*, 1, \Sigma, \Pi, \Pi^*)\)-type-category''.
    Section 3.5 tells us that \(\B\) is such a category.
\end{proof}

Next, we specialize to sheaf toposes.

@@

Finally, we set the site to a syntactic site, to understand classifying toposes.

@@

    \vspace{1em}

    \noindent The syntactic category is guaranteed to be finitely complete, with a canonical construction for pullbacks.
    Specifically, the empty context is terminal,
    and the pullback of the span \(\Gamma \xrightarrow{f} \Xi \xleftarrow{g} \Delta\) is the context \((\gamma : \Gamma, \delta : \Delta, f\gamma = g\delta)\).
